---
title: "QUESTION 1"
author: "Riya Sebastian"
date: "2024-06-12"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading necessary packages
```{r packages}
pacman::p_load(tidyverse,tidymodels,bookdown,ggplot2)
```

# Reading the data to a tibble
```{r data}
board_game=readRDS("board_game.rds")
board_game=tibble(board_game)
board_game
```

# a. Selecting the variables
```{r a}
board_game = board_game %>% select (
  primary, year, boardgamemechanic, minplaytime, maxplaytime, average
  )
board_game 
```

# b. Renaming average 
```{r b}
board_game = board_game %>% rename (rating = average)
board_game
```

# c. Removing games released before 2016 or after 2020
```{r c}
board_game = board_game[board_game$year >= 2016 & board_game$year <= 2020, ]
board_game
```


# d. Creating new variable "duration" and filtering out data with duration = 0
```{r d}
board_game = board_game %>% mutate (duration = sqrt(minplaytime * maxplaytime))
board_game = board_game %>% filter(duration!=0)
board_game
```

# e. Creating a variable called dice
```{r e}
board_game = board_game %>%
  mutate(dice =case_when(
    str_detect(boardgamemechanic, "Dice") ~ "Dice game",
    is.na (boardgamemechanic) ~"Not a dice game",
    TRUE ~ "Not a dice game"))
board_game
```

# f. Creating a variable called rating_7.5
```{r f}
board_game = board_game %>% mutate(rating_7.5 = ifelse(rating >= 7.5, TRUE, FALSE))
board_game
```

# g. Creating a box plot
```{r g, fig.cap = 'Note: Games without a mechanic have been classified as "Not a dice game".',fig.align='center'}
board_game %>% 
  ggplot(aes(x = factor(year), y = duration, fill = factor(rating_7.5))) + 
  scale_fill_manual(values = c("FALSE" = "yellow", "TRUE" = "#1b9e77")) + 
  geom_boxplot() +
  scale_y_continuous(trans = 'log10', breaks = c(1, 10, 100, 1000, 10000))  +
  labs(x= "Year", y="Duration (minutes)",fill = "Rating >= 7.5",title=
         "Game Duration by Year", subtitle = 
         "Colour indicates if game have a rating of 7.5 or above") + 
  facet_wrap(~dice,ncol=1) + theme(legend.position = "bottom")

```
Figure \@ref(fig:g) Side-by-side boxplot of game duration (in minutes) by year, separated by whether or not each game is a dice game, and whether or not each game has a rating of at least 7.5.


